Main Analyses of the Combined Dataset

### 2021-05-11 manuscript/manuscript.tex
more revisions

Notably, this relationship has few outliers and is relatively homoscedastic.
This is in part because of the log-log scale, in contrast to previous work
\parencite{Nei1984-zi,Soule1976-he}; see Supplementary Figure
\ref{suppfig:figure-1-linear} for a version on a log-linear scale. However, it
is noteworthy that few taxa have diversity estimates below $10^{-3.5}$
differences per basepair. Those that do, lynx (\emph{Lynx Lynx}), wolverine
(\emph{Gulo gulo}), and Massasauga rattlesnake (\emph{Sistrurus catenatus})
face habitat fragmentation and declining population sizes. These three species
are all in the IUCN Red List, but are listed as least concern (though their
presence in the Red List indicates they are of conservation interest). I also
investigated the relationship between species' IUCN Red List categories (an
ordinal scale of how threatened a species is) and both diversity and population
size, finding that species categorized as more threatened have both smaller
population sizes and reduced diversity, compared to non-threatened species
(Figure \ref{fig:figure-1-redlist}) consistent with past work
\parencite{Spielman2004-mt}. A linear model of diversity regressed on
population size has lower AIC when the IUCN Red List categories are included,
and the estimates of the effect of IUCN status are all negative on diversity,
though not all are significant in part because some categories have three or
fewer species (Supplementary Table \ref{supptable:tbl-1}).

  \label{fig:figure-1-redlist}
\end{figure}

@@

\section{Diversity and IUCN Red List Status}
\begin{table}[ht]
  \centering
  \begin{tabular}{rrrr}
    \hline
 & mean & 2.5 \% & 97.5 \% \\ 
 \hline
    $\beta_0$ & -2.70 & -3.00 & -2.40 \\ 
    $\beta_{LC}$ & -0.37 & -0.55 & -0.19 \\ 
    $\beta_{NT}$ & -0.19 & -0.80 & 0.41 \\ 
    $\beta_{VU}$ & -0.31 & -0.81 & 0.19 \\ 
    $\beta_{EN}$ & -0.39 & -0.71 & -0.06 \\ 
    $\beta_{CR}$ & -0.03 & -0.65 & 0.59 \\ 
    $\beta_{N_c}$ & 0.06 & 0.04 & 0.09 \\ 
    \hline
  \end{tabular}

>>>>

Notably, this relationship has few outliers and is relatively homoscedastic.
This is in part because of the log-log scale, in contrast to previous work
\parencite{Nei1984-zi,Soule1976-he}; see Supplementary Figure
\ref{suppfig:figure-1-linear} for a version on a log-linear scale. However, it
is noteworthy that few taxa have diversity estimates below $10^{-3.5}$
differences per basepair. Those that do, lynx (\emph{Lynx Lynx}), wolverine
(\emph{Gulo gulo}), and Massasauga rattlesnake (\emph{Sistrurus catenatus})
face habitat fragmentation and declining population sizes. These three species
are all in the IUCN Red List, but are listed as least concern (though their
presence in the Red List indicates they are of conservation interest). In
Supplementary Material Section \namref{supinfo:div-iucn}, I explore more about
the relationships between IUCN Red List status, diversity, and population size.

@@

\section{Diversity and IUCN Red List Status}
\label{supinfo:div-iucn}

I also investigated the relationship between species' IUCN Red List categories
(an ordinal scale of how threatened a species is) and both diversity and
population size, finding that species categorized as more threatened have both
smaller population sizes and reduced diversity, compared to non-threatened
species (Supplementary Figure \ref{suppfig:figure-1-redlist}) consistent with
past work \parencite{Spielman2004-mt}. A linear model of diversity regressed on
population size has lower AIC when the IUCN Red List categories are included,
and the estimates of the effect of IUCN status are all negative on diversity,
though not all are significant in part because some categories have three or
fewer species (Supplementary Table \ref{supptable:tbl-1}).

  \label{suppfig:figure-1-redlist}
\end{figure}

\begin{table}[ht]
  \centering
  \begin{tabular}{rrrr}
    \hline
 & mean & 2.5 \% & 97.5 \% \\ 
 \hline
    $\beta_0$ & -2.70 & -3.00 & -2.40 \\ 
    $\beta_{LC}$ & -0.37 & -0.55 & -0.19 \\ 
    $\beta_{NT}$ & -0.19 & -0.80 & 0.41 \\ 
    $\beta_{VU}$ & -0.31 & -0.81 & 0.19 \\ 
    $\beta_{EN}$ & -0.39 & -0.71 & -0.06 \\ 
    $\beta_{CR}$ & -0.03 & -0.65 & 0.59 \\ 
    $\beta_{N_c}$ & 0.06 & 0.04 & 0.09 \\ 
    \hline
  \end{tabular}

### 2021-05-11 notebooks/main_analysis.Rmd
more revisions

# data for the log-normal fit
log10_popsize_seq <- seq(5, 18, length.out=100)
ml_dat <- list(N = nrow(dt_ml_ns), 
              log10_popsize = dt_ml_ns$log10_popsize,
              map_length = dt_ml_ns$map_length, 
              cor = ml_cor_ns,
              x_new = log10_popsize_seq)

@@

# this is the one used in figures -- assumes Ne = Nc
# this got a bit unwieldy as I tried different things... oh well.
dml_full <- da_ml %>%
     # estimate from Elyashiv et al.
     # mutate(J22 = 100*J22s(dm_seldist$s,  dm_seldist$g, 10^log10_popsize, Inf)) %>%
     mutate(J22 = dm_J22_Elyashiv2) %>%
     mutate(Ne_N_HK95 = HK95(map_length, U=dm_U)) %>% 
     mutate(Ne_N_HK95_strongBGS = HK95(map_length, U=10*dm_U)) %>% 
     # both BGS and RHH
     mutate(Ne_N_RHH_BGS = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22)) %>%
     mutate(Ne_N_RHH_BGS_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                             B=Ne_N_HK95, gamma=dm_gamma, 
                                             J22=J22 * 10)) %>%

@@

     # 100-folder greater Nc
     mutate(Ne_N_RHH_BGS3 = RHH_BGS(1, L=map_length, N=100*10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=dm_J22_Elyashiv))

@@

@@

plot(dml$log10_popsize, dml$log10_diversity, pch=19, 
                  xlim=c(5, 18),
                  ylim=c(-3, 1))

>>>>

# data for the log-normal fit
log10_popsize_seq <- seq(5, 15.2, length.out=100)
ml_dat <- list(N = nrow(dt_ml_ns), 
              log10_popsize = dt_ml_ns$log10_popsize,
              map_length = dt_ml_ns$map_length, 
              cor = ml_cor_ns,
              x_new = log10_popsize_seq)

@@

# this is the one used in figures -- assumes Ne = Nc
# this got a bit unwieldy as I tried different things... oh well.
dml_full <- da_ml %>%
     # estimate from Elyashiv et al.
     # mutate(J22 = 100*J22s(dm_seldist$s,  dm_seldist$g, 10^log10_popsize, Inf)) %>%
     mutate(J22 = dm_J22_Elyashiv2) %>%
     mutate(Ne_N_HK95 = HK95(map_length, U=dm_U)) %>% 
     mutate(Ne_N_HK95_strongBGS = HK95(map_length, U=10*dm_U)) %>% 
     # RHH for comparison
     mutate(Ne_N_RHH = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                   B=1, gamma=dm_gamma, 
                                   J22=J22)) %>%
     # both BGS and RHH
     mutate(Ne_N_RHH_BGS = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22)) %>%
     mutate(Ne_N_RHH_BGS_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                             B=Ne_N_HK95, gamma=dm_gamma, 
                                             J22=J22 * 10)) %>%

@@

     # 100-folder greater Nc
     mutate(Ne_N_RHH_BGS3 = RHH_BGS(1, L=map_length, N=100*10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22))

@@

What are the relative strengths of BGS and RHH?

dml_full %>% mutate(ratio = log10(Ne_N_RHH / Ne_N_HK95)) %>% 
  ggplot(aes(log10_popsize, ratio)) + geom_point()

@@

plot(dml$log10_popsize, dml$log10_diversity, pch=19, 
                  xlim=c(5, 16),
                  ylim=c(-3, 1))
